### 2017-01-10 conclusion.tex
Conlusion with Future Work

% Chapter 5: Conclusion

\chapter{Conclusions and Future Work}
\label{sec:conclusion}

Plasma irregularities in the F-region ionosphere have been observed for decades.  However, despite extensive data sets collected over these years, supprisingly little is known about these irregularities, especially at small scales.  This is in part due to the complex and turbulent nature of the ionosphere, particularly in the polar regions, but also due to some of the inherent limitations in traditional observation techniques which can make it very challenging to identify exactly what processes are operational.  In this Chapter, we summarize the results that have previously been discussed in Chapters \ref{sec:paper1}--\ref{sec:paper3} and consider what overarching conclusions can be derived from them which further our understanding of small-scale polar F-region plasma irregularities.

\section{Summary and Conclusions}
\label{sec:summary}

\section{Future Work}
\label{sec:futurework}

Although much progress has been made in understanding small-scale plasma structuring and HF radar observations, many of the results presented in this thesis give rise to further questions that must be addressed.

In Chapter \ref{sec:paper1}, the occurence of radar backscatter durring nighttime in the polar cap was found to be much higher than expected, particularly given that raytracing simulations predict that the orthagonality condition is rarely met and therefore there should be little to no backscatter.  The proposed hypothesis to explain this discreperency is that polar patches, which are not typically well represented by ionospheric density models, occur frequently in the nighttime F region and, on average, enhance the background plasma density enough that propagation conditions are sufficient for HF radar backscatter to regularly be observed.  Experimental surveys of polar patches have shown low occurrence at nighttimes, however the occurrence of patches and their intensity increase at equinox, which is where frequent nighttime backscatter is predominantly observed.  The expected signature in the radar data from this kind of senario would be a night with a series of isolate ``patches'' of backscatter.  An example of such an event was presented, Figure \ref{whatever the hell THIS'll be}, but it is unknown how often events like this occur.  Additionally, the intensity and size of a patch nessesary to cause the radar beam to refract sufficiently must be established, possibly through further use of raytracing simulations with customizable density profiles.

One of the major drawbacks to linear GDI theory is its limitation of applicability only to long wavelengths, which means if the F region, it cannot be directly applicable to decameter-scale irregularities measured by HF radars.  Generally, it is assumed that some form of turbulent cascade connects plasma structures of different scale sizes in the ionosphere such that predictions made at large scales can still be applied to decameter-scale observations.  This helps explain some of the observations of asymmetry around polar patches.  This asymmetry was also confirmed in the GDI growth rate from the results shown in Chapter \ref{sec:paper2}.  However, these results also predicted anisotropy in the growth rate, which is more challenging to confirm from small-scale irregularity obsersavtions, although some evidence of such has been provided in Chapter \ref{sec:paper3}.  It may be possible to further the theoretical understanding of GDI at small scales by developing a three dimentional picture of it using kinetic theory.  \citet{Basu1995} developed such a formulation for a specific geometry throughout all altitudes within the ionosphere, but expanding this to general vector directions would most likely be extremely mathematically intensive.

The results presented in Chapter \ref{sec:paper3} could also be improved with a more extensive data set.  The event chosen for analysis in that study was in 2011 and was selected on the basis of good data coverage from both RISR-N and RKN and a strong density gradient being apparent in the overlapping FoV.  Since then, RISR-C has been deployed with a FoV just south of RISR-N (Figure \ref{fig:isrmap}) such that observations between both radars can be coordinated and they can act as a single instrument with a FoV twice as large.  This allows a polar patch to be tracked over a longer distance and more individual data points to be captured.  Additionally, specific modes can be utilized that give a much finner grid of RISR beams, allowing measurments to be made closer together and smaller scale gradients to be be considered.  The downside to these modes is that it takes longer for each set of measurements to be made, so the temporal cadence decreases.  However, the required integration time for an ISR decreases if the background ionospheric density is higher, so the the temporal resolution can improve for measurments taken durring summer or daytime.  Another advantange to the RISR-N/RISR-C locations is that the SuperDARN radars at Clyde River (CLY) and Inuvik (INV) (Figure \ref{fig:superdarnmap}), both of which share at least part of the RISR-N/RISR-C FoV, but observe it from different directions.  This presents an unique oppertunity to further investigate the directional dependence of small-scale irregularities by observing the differences in backscatter observed from the same scattering volume at different directions.  Although straight forward in theory, this experiment would most likely prove nontrivial in practice.  Backscatter tends to be limited to the statistical high occurrence band discussed in Chapter \ref{sec:paper1}, and at least with INV, this band probably occurs at closer ranges than where the FoV overlaps other radars, so coincident backscatter is probably rare.  The probability of backscatter being observed is better when the background density is within the range where propogation conditions are most favorable, when the ionosphere is sunlit.  Additionally, it is adventagious to choose a time when the occurrence of backscatter is statistically high.  Based on results from Chapter \ref{sec:paper1}, the best time for such an experiment would therefore be daytime durring an equinox month.