### 2017-02-01 acronyms.tex
List of Acronyms added

% Acronyms

\begin{longtable}{ll}
\textbf{AACGM} & Altitude-Adjusted Corrected Geomagnetic coordinate system \\
\textbf{AC} & Alternating Code Pulse \\
\textbf{ACF} & Autocorrelation Function \\
\textbf{AMISR} & Advanced Modular Incoherent Scatter Radar \\
\textbf{ASI} & All-Sky Imager \\
\textbf{ASIP} & All Sky Imaging Photometer \\
\textbf{BSN} & Bow Shock Nose \\
\textbf{CCI} & Current-Conductive Instability \\
\textbf{CCMC} & Community Coordinated Modeling Center \\
\textbf{CLY} & Clyde River SuperDARN Radar \\
\textbf{CME} & Coronal Mass Ejection \\
\textbf{COSPAR} & Committee on Space Research \\
\textbf{CSR} & Coherent Scatter Radar \\
\textbf{EISCAT} & European Incoherent Scatter Association \\
\textbf{ESR} & European Scatter Radar \\
\textbf{FAI} & Field Aligned Irregularity \\
\textbf{FBI} & Farley-Buneman Instability \\
\textbf{FoV} & Field of View \\
\textbf{GDI} & Gradient Drift Instability \\
\textbf{GNSS} & Global Navigation Satellite System \\
\textbf{GPS} & Global Positioning System \\
\textbf{HF} & High Frequency \\
\textbf{IMF} & Interplanetary Magnetic Field \\
\textbf{INV} & Inuvik SuperDARN radar \\
\textbf{IRI} & International Reference Ionosphere \\
\textbf{ISR} & Incoherent Scatter Radar \\
\textbf{KHI} & Kelvin-Helmholtz Instability \\
\textbf{LoS} & Line of Sight \\
\textbf{LP} & Long Pulse \\
\textbf{MCM} & McMurdo Station SuperDARN radar \\
\textbf{MLAT} & Magnetic Latitude \\
\textbf{MLON} & Magnetic Longitude \\
\textbf{MSISE} & Mass Spectrometer - Incoherent Scatter Extended \\
\textbf{PFISR} & Poker Flat Incoherent Scatter Radar \\
\textbf{PCS} & Polar Cap South index \\
\textbf{RISR} & Resolute Bay Incoherent Scatter Radar \\
\textbf{RISR-C} & Resolute Bay Incoherent Scatter Radar Canada \\
\textbf{RISR-N} & Resolute Bay Incoherent Scatter Radar North \\
\textbf{RKN} & Rankin Inlet SuperDARN radar \\
\textbf{SFJ} & Sondrestrom \\
\textbf{STARE} & Scandanavian Twin Auroral Radar Experiment \\
\textbf{SuperDARN} & Super Dual Auroral Radar Network \\
\textbf{SZA} & Solar Zenith Angle \\
\textbf{TEC} & Total Electron Content \\
\textbf{URSI} & International Union of Radio Science \\
\textbf{UT} & Universal Time \\
\textbf{VHF} & Very High Frequency
\end{longtable}