### 2017-02-02 acknowledgements.tex

% Acknowlegements

First and foremost, I would like to thank my advisor, Dr. Roman Makarevich for his guidance throughout this project, which would not have been possible without his support, feedback, and constructive criticism.  I would also like to acknowledge the efforts of the other members of my committee, Dr. William Bristow, Dr. Mark Conde, and Dr. Hui Zhang for always providing new ideas and perspectives.  Additionally, I truely appreciate the support and encouragement I received from the entire physics department at UAF, particularly Dr. David Neuman, Dr. Renate Wackerbauer, Dr. Curt Szuberla, Dr. Chung-Sang Ng, Dr. Peter Delamere, Ellen Craig and Jeanie Talbot.  I would like to extend a particularly special thank you to Flora G**** and all the staff of the Keith B. Mather Libray for providing invaluable assistance in retrieving sources for my research, no matter how old and obscure my requests were.

To my family, thank you all for the supporting me for years, particularly my Mom, for making me a much better writer than I have any right to be as a scientist, my Dad, for always being there to answer my programming questions and helping me fix my code when I had now idea how I had broken it, and my brother, Henry, generally for just existing and providing a friendly voice of reason when I really needed it.  Additionally, thank you to my friends Katie, Ari, Bonnie, Heather, Jayne, and Stephanie for being there over these last few years despite being 4000 miles away, to Laura for giving me perspective and a laugh regardless of the time difference.  Finally, I would like to extend a very special thank you to Allison, Rosie, Emma, and Kinley on the homefront for their tireless effort in keeping me sane, healthy, and happy throughout this whole process.  I couldn't have done it without you guys.  Thank you.